---
title: "Question3_2"
output: pdf_document
author: "Hukai Luo"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract:
    In this assignment, we use the probability density function to find the log-likelihood function and the method-of-moments estimator of $\theta$. Then we will discuss the influence of different initial value when using Newton-Raphson method to estimate the Maximum Likelihood Estimation for $\theta$.
---

# 1 Many local maxima

The probability density function with parameter $\theta$:
\begin{align*}
  f(x; \theta) = \frac{1 - \cos(x - \theta)}{2\pi},   
  \quad 0\le x\le 2\pi, \quad \theta \in (-\pi, \pi).
\end{align*}
And the random sample from the distribution is

$x_i = 3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96, 2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52$

- Find the the log-likelihood function of $\theta$ based on the sample and plot it between $-\pi$ and $\pi$.
- Find the method-of-moments estimator of $\theta$.  That is, the estimator $\tilde\theta_n$ is value of $\theta$ with
\begin{align*}
  E (X \mid \theta) = \bar X_n,
\end{align*}
where $\bar X_n$ is the sample mean. This means you have to first find the expression for $E (X \mid \theta)$.
- Find the MLE for $\theta$ using the Newton--Raphson method initial value $\theta_0 = \tilde\theta_n$.
- What solutions do you find when you start at 
$\theta_0 = -2.7$ and $\theta_0 = 2.7$?
- Repeat the above using 200 equally spaced starting values
between $-\pi$ and $\pi$.  Partition the values into sets of attraction. That is, divide the set of starting values into separate groups, with each group corresponding to a separate unique outcome of the optimization.

# 2 Log-likelihood function

The log-likelihood function with parameter $\theta$ and $x$:
\begin{align*}
 \ell(\theta) = \sum^n_{i=1}{ln(1-\cos{(x-\theta)})} - nln2\pi
\end{align*}
\newline
then plot it between $-\pi$ and $\pi$:
```{r code, results='hide'}
X = c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96,
      2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52)
logfunction <- function(x) {                        #calculate the log likelihood function
  logfunction <- 0
  for (i in 1:length(X)){
    logfunction <- logfunction+log(1-cos(X[i]-x))-log(2*pi)
  }
  logfunction
}

library("ggplot2")                                        #plot the function against theta
ggplot(data.frame(x= c(-pi,pi)),aes(x = x))+
         stat_function(fun = function(x) logfunction(x))+
         labs(x=expression(~theta),y="log likelihood function")
```

# 3 Method-of-moments estimator
The expression for $E (X \mid \theta)$ is:
\begin{align*}
E(X|\theta) & =\int^{2\pi}_{0}x\frac{1-\cos(x-\theta)}{2\pi}dx\\
            & =\frac{1}{2\pi}\left[\int^{2\pi}_{0}xdx-\int^{2\pi}_{0}x\cos(x-\theta)dx\right]\\
            & =\frac{1}{2\pi}\left[\frac{1}{2}x^2\big |^{2\pi}_{0}-xsin(x-\theta)\big |^{2\pi}_{0}\right]\\
            & =\frac{1}{2\pi}\left(2\pi^{2}+2\pi\sin\theta\right)\\
            & =\pi+\sin\theta
\end{align*}
\newline \newline
And $\bar X_n$ is 3.236842, so we can get $\theta$:
\begin{align*}
\theta & = \arcsin(mean(x)-\pi)\quad or\quad \pi-\arcsin(mean(x)-\pi)\\
       & = 0.095394\quad or\quad 3.046199\\
\end{align*}

# 4 Newton--Raphson method

Find the MLE for $\theta$ using the Newton--Raphson method initial value $\theta_0 = \tilde\theta_n$. In this question, we will calculate$\ell'(\theta),\ell''(\theta)$ as follow:
\begin{align*}
\ell'(\theta) = \sum^n_{i=1}{\frac{-sin(x-\theta)}{1-cos(x-\theta)}}\\
\ell"(\theta) = \sum^n_{i=1}{\frac{1}{cos(x-\theta)-1}}
\end{align*}
Newton--Raphson method:
```{r code2, results='hold'}
library(pracma)
derivative_1 <- function(x) {         #calculate the derivative of log likelihood function
  dev1 <- 0
  for (i in 1:length(X)){
    dev1 <- dev1-sin(X[i]-x)/(1-cos(X[i]-x))
  }
  dev1
}
derivative_2 <- function(x) {  #calculate the second derivative of log likelihood function
  dev2 <- 0
  for (i in 1:length(X)){
    dev2 <- dev2+1/(cos(X[i]-x)-1)
  }
  dev2
}
x_initial <- 0.095394
result1 <- newtonRaphson(derivative_1,x_initial,dfun = derivative_2)
x_initial <- 3.046199
result2 <- newtonRaphson(derivative_1,x_initial,dfun = derivative_2)
```
If we start with $\theta = 0.095394$ we will get $root = 0.003118157$, when we start with $\theta = 3.046199$ the root is $3.170715$.

# 5 Solutions find  using selected $\theta = \pm2.7$

We can just change the initial vaule of $\theta$ to find the root:
```{r code3, results='hold'}
x_initial <- 2.7
result3 <- newtonRaphson(derivative_1,x_initial,dfun = derivative_2)
x_initial <- -2.7
result4 <- newtonRaphson(derivative_1,x_initial,dfun = derivative_2)
```
If we start with $\theta = 2.7$ we will get $root = 2.848415$, when we start with $\theta = -2.7$ the root is $-2.668857$.

# 6 Solutions using $\theta$ between $-\pi$ and $\pi$

```{r code4, results='hold',warning=FALSE}
library("pander")
options(digits = 5)
x_initial <- seq(-pi,pi,length = 200)
result <- matrix(0,1,length(x_initial))
for (i in 1:length(x_initial)){
  result[i] <- newtonRaphson(derivative_1,x_initial[i],dfun = derivative_2) 
}
plot(x_initial, result, title("Initial values and roots"))
table = rbind(x_initial,result)
set.caption("Initial values and roots")
pander(table,split.table = 100,style = 'rmarkdown')
```
So we have get the table of initial value and root, we can see there are several groups of the results, they are 
```{r code5, include=FALSE}
roots <- c(-3.1125,-2.7866,-2.6689,-2.5094,-2.3883,-2.2979,-2.2322,-1.6627,-1.4475,-0.9544,-0.0031182,-0.81264,2.0072,2.237,2.3747,2.4884,2.8484,3.1707)
freq <- c(11,2,5,6,1,4,1,24,1,19,42,46,8,2,6,2,15,5)
table1 <- rbind(roots,freq)
```
```{r code6, results='hold'}
pander(table1)
```














